Automatic Bicycle Balance Assistance Reduces Probability of Falling at Low Speeds When Subjected to Handlebar Perturbations\\Version 4.dev

### 2025-10-21 main.tex
Added the mean age of the particpants (does not include 4 missing ages in the data files).
Now includes two eigenvalue figures.

\subsection{Balance Assist Control}
%
We use a forward speed \(v\) gain scheduled proportional roll angular rate
feedback controller to stabilize the bicycle. In the speed range tested, the
commanded steer torque \(T^\textrm{m}_\delta\) from the steer motor follows the
control law
%
\begin{align}
  T^\textrm{m}_\delta = -k_{\dot{\phi}}\dot{\phi} = \kappa(v_{stable} - v)\dot{\phi}
  \label{eq:implemented-controller}
\end{align}
%
where \(v_{\textrm{stable}} = 4.7~\si{\meter\per\second}\) is approximately the
average stable speed predicted from the open loop bicycle rigid-rider dynamics.
We use the gain factor \(\kappa=8\) (low) and \(\kappa=10\) (high) to tune the gain
magnitude during the experiments. Scaling the proportional feedback gain
linearly with respect to speed stabilizes the normally unstable weave mode of
the bicycle down to 2.9~\si{\kilo\meter\per\hour} for the riderless bicycle and
4.6~\si{\kilo\meter\per\hour} for the ridden bicycle as shown in
Figure~\ref{fig:eigenvalues}.
%
\begin{figure}
  \centering
  \includegraphics[width=160mm]{figures/balance-assist-eig-vs-speeds.png}
  \caption{Uncontrolled (upper row) and controlled with \(\kappa=10\) (lower
    row) root locus of the eigenvalue components (real: solid, imaginary:
    dashed) plotted versus speed for the linear Carvallo-Whipple bicycle model
    without the inertial effect of a rigid rider (left) and with a rigid rider
    (right). Vertical dotted and dotted-dashed lines indicate the two speeds we
    perform experiments at: 6~\si{\kph} and 10~\si{\kph}. The grey shaded
    region is the linear stable speed range. Geometric and inertial parameters
    for these plots were estimated with the methods presented in
    \citet{Moore2012} and software packages
    BicycleParameters~1.1.1~\citep{Moore2024} and
    Yeadon~1.5.0~\citep{Dembia2015} and are shown in
    Appendix~\ref{sec:physical-parameters}.}
  \label{fig:eigenvalues}
\end{figure}

@@

\subsection{Protocol}
%
We recruited 26 able-bodied young adults (20-36 years old) to participate in
the experiments. The participants were all confident in their cycling skills
and had cycled at least once in the last month. All participants consented to
the experiment and could decline to continue at any time. The study was
approved by Delft University of Technology's Human Research Ethics Council
(record \#3897).

>>>>

\subsection{Balance Assist Control}
%
We use a forward speed \(v\) gain scheduled proportional roll angular rate
feedback controller to stabilize the bicycle. In the speed range tested, the
commanded steer torque \(T^\textrm{m}_\delta\) from the steer motor follows the
control law
%
\begin{align}
  T^\textrm{m}_\delta = -k_{\dot{\phi}}\dot{\phi} = \kappa(v_{stable} - v)\dot{\phi}
  \label{eq:implemented-controller}
\end{align}
%
where \(v_{\textrm{stable}} = 4.7~\si{\meter\per\second}\) is approximately the
average stable speed predicted from the open loop bicycle rigid-rider dynamics.
We use the gain factor \(\kappa=8\) (low) and \(\kappa=10\) (high) to tune the gain
magnitude during the experiments. Scaling the proportional feedback gain
linearly with respect to speed stabilizes the normally unstable weave mode of
the bicycle down to 2.9~\si{\kilo\meter\per\hour} for the riderless bicycle and
4.6~\si{\kilo\meter\per\hour} for the ridden bicycle as shown in
Figure~\ref{fig:eigenvalues}.
%
\begin{figure}
  \centering
  \includegraphics[width=160mm]{figures/balance-assist-gain-8-eig-vs-speeds.png}
  \caption{Uncontrolled (upper row) and controlled with \(\kappa=8\) (lower
    row) root locus of the eigenvalue components (real: solid, imaginary:
    dashed) plotted versus speed for the linear Carvallo-Whipple bicycle model
    without the inertial effect of a rigid rider (left) and with a rigid rider
    (right). Vertical dotted and dotted-dashed lines indicate the two speeds we
    perform experiments at: 6~\si{\kph} and 10~\si{\kph}. The grey shaded
    region is the linear stable speed range. Geometric and inertial parameters
    for these plots were estimated with the methods presented in
    \citet{Moore2012} and software packages
    BicycleParameters~1.1.1~\citep{Moore2024} and
    Yeadon~1.5.0~\citep{Dembia2015} and are shown in
    Appendix~\ref{sec:physical-parameters}.}
  \label{fig:eigenvalues}
\end{figure}

\begin{figure}
  \centering
  \includegraphics[width=160mm]{figures/balance-assist-gain-10-eig-vs-speeds.png}
  \caption{Uncontrolled (upper row) and controlled with \(\kappa=10\) (lower
    row) root locus of the eigenvalue components (real: solid, imaginary:
    dashed) plotted versus speed for the linear Carvallo-Whipple bicycle model
    without the inertial effect of a rigid rider (left) and with a rigid rider
    (right). Vertical dotted and dotted-dashed lines indicate the two speeds we
    perform experiments at: 6~\si{\kph} and 10~\si{\kph}. The grey shaded
    region is the linear stable speed range. Geometric and inertial parameters
    for these plots were estimated with the methods presented in
    \citet{Moore2012} and software packages
    BicycleParameters~1.1.1~\citep{Moore2024} and
    Yeadon~1.5.0~\citep{Dembia2015} and are shown in
    Appendix~\ref{sec:physical-parameters}.}
  \label{fig:eigenvalues}
\end{figure}

@@

\subsection{Protocol}
%
We recruited 26 able-bodied young adults between 20 and 36 years old (mean 26
years) to participate in the experiments. The participants were all confident
in their cycling skills and had cycled at least once in the last month. All
participants consented to the experiment and could decline to continue at any
time. The study was approved by Delft University of Technology's Human Research
Ethics Council (record \#3897).
